---
title: "Imputed Analyses"
author: "Berenice Anaya"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Installing packages 
list.of.packages <- c("dplyr", "plyr","stats","ggplot2", "psych", "nlme","tinytex","lattice", "mitml", "lme4","interactions", "jtools", "cowplot", "tidyr", "readr", "mice", "broom", "miceadds", "broom.mixed", "pander", "ggsignif", "here", "stringr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages) 
# Loading packages
lapply(list.of.packages, require, character.only = TRUE)
```

# Introduction
This script runs the analyses for the paper with multiple imputations to probe sensitivity/ robustness of the results

# 1. Loading Datasets

## 1.1. Behavioral Datasets
```{r cars}
demo <- read.csv(here("data", "demo.csv"), header = T)
demo <- demo%>%
  dplyr::select(modid, screen_delivery_ga_weeks:child_sex, mom_age_risk, mat_med_risk_sum, child_age_y1_assessment)%>%
  dplyr::rename(id = modid)

ibq <- read.csv(here("data","Anaya_02_20_2024_ibqr.csv"), header=T)

ibq <- ibq %>%
    tidyr::drop_na(ibqr_neg) %>%
  dplyr::group_by(modid) %>%
  dplyr::summarise(
    ibq_neg_m = mean(ibqr_neg, na.rm = TRUE),
    n_obs = sum(!is.na(ibqr_neg)),
    .groups = "drop"
  )%>%
  dplyr::rename(id = modid)

  
pren_stress <- read.csv(here("data","Stress_preg.csv"), header = T)
colnames(pren_stress) <- c("id", "dep_1", "dep_2", "dep_3", "dep_4", "stress_1", "stress_2", "stress_3", "stress_4")

ses <- read.csv(here("data","social_dis.csv"), header = T)
ses <- ses%>%
  dplyr::select(MODID, disadv_prenatal)%>% #selecting only prenatal disadvantage
  dplyr::rename(id = MODID)

exc <- read.csv(here("data","exclusion.csv"), header = T)
exc <- exc%>%
  dplyr::select(modid, combined_exclusion_initial_analyses, irb_exclusion)%>%
  dplyr::rename(id = modid)%>%
  dplyr::mutate(exc_all = if_else(combined_exclusion_initial_analyses == 1 | irb_exclusion == 1, 1, 0))%>%
  dplyr::select(id, exc_all)

# 81 subjects were excluded - running imputations on all analyses then, including growth models of caregiver stress

QC <- read.csv(here("data","HAPPE_pipelineQC_Y1.csv"), header = T) # need this dataset
QC$id <- substr(QC$Row, 1,7)
QC <- QC%>%
  dplyr::select(id, r.alldata.post.pre.waveleting)%>%
  dplyr::mutate(exc = if_else(r.alldata.post.pre.waveleting < .15, 1, 0)) # setting exclusion to < .15
```


## 1.2. Delta-Beta Synchrony Datasets (different datasets per trial type)
```{r}
# 168 subjects with EEG after HAPPE
odd_sync <- read.csv(here("data","OddSynchrony.csv"), header = T)
std_sync <- read.csv(here("data","StdrSynchrony.csv"), header = T)
aft_sync <- read.csv(here("data","AfterSynchrony.csv"), header = T)

# renaming columns and merging wide
names(odd_sync)[2:5] <- paste0(names(odd_sync)[2:5], "_odd")
names(std_sync)[2:5] <- paste0(names(std_sync)[2:5], "_stdr")
names(aft_sync)[2:5] <- paste0(names(aft_sync)[2:5], "_after")

wide_all_sync <- merge(odd_sync, std_sync, by = "id")
wide_all_sync <- merge(wide_all_sync, aft_sync, by = "id")
wide_all_sync <- wide_all_sync%>%
  dplyr::select(-c(trial_odd, trial_after, trial_stdr))

# excluding data based on HAPPY QC
wide_all_sync <- merge(wide_all_sync, QC, by="id", all = T)
wide_all_sync <- wide_all_sync%>%
  dplyr::mutate(across(
    -id,
    ~ if_else(exc == 1, NA, .)
  ))%>%
  dplyr::select( -exc)

describe(wide_all_sync)
# 124 participants with usable delta-beta coupling scores


# shaping to long
all_sync <- reshape(wide_all_sync,
                    varying = 2:10,
                    timevar = "trial",
                    idvar = "id",
                    direction = "long",
                    sep = "_")

```

## 1.3. Visualizing correlations between coupling scores between trial types
```{r}
library(ggcorrplot)

vars <- wide_all_sync %>% 
  select(-r.alldata.post.pre.waveleting)%>%
  select(where(is.numeric))

corr <- cor(vars, use = "pairwise.complete.obs")

p.mat <- cor_pmat(vars)

ggcorrplot(
  corr,
  p.mat = p.mat,
  type = "lower",
  lab = TRUE,
  insig = "blank"
)

```


# 2. Missingness Analysis
```{r}
library(mice)
library(naniar)
```

## 2.1. Quantifying missingness
We first merge all datasets 
```{r}
# we first merge all relevant study variables
analysis_df_all <- merge(pren_stress, ibq, by ="id", all = T)
analysis_df_all <- merge(analysis_df_all, wide_all_sync, by ="id", all = T)

# excluding participants from the analytical sample  based on global exclusion criteria
analysis_df_all <- merge(analysis_df_all, exc, by="id", all = T)
analysis_df_all <- analysis_df_all %>%
  dplyr::mutate(across(
    -id,
    ~ if_else(exc_all == 1, NA, .)
  ))%>%
  dplyr::select(-exc_all)

# now merging with demo and ses to get information on missing participants 
df_all_missing <- merge(analysis_df_all, ses, by="id", all = T)
df_all_missing <- merge(demo, df_all_missing, by="id", all = T)


describe(df_all_missing)

# identifying rows with all NAs
rows_all_na <- df_all_missing %>%
  filter(if_all(-c(id,disadv_prenatal),  is.na))

# Wrangling for Mplus analysis
mplus_elabe_df <- df_all

mplus_elabe_df <- mplus_elabe_df%>%
  #mutating id column for mplus
  dplyr::mutate(id_imp = as.integer(str_extract(id, "\\d+")))%>%
  #setting NA values to -999
  dplyr::mutate(across(everything(), ~ replace_na(., 999)))

# Exporting
write.table(mplus_elabe_df, "mplus_df_names.csv", sep=",", row.names = F, na = "999", quote=F)

# creating new dataset for MICE imputations
df_all<- df_all_missing
```


### Missing numbers
```{r}
# missing pattern all
md.pattern(df_all)

# missing pattern repeated variables
df_all %>%
  select(id, dep_1:stress_4) %>%
  md.pattern()

```

## 2.2. Visualizing co-occurring missing data
```{r}
# all 
vis_miss(df_all)
gg_miss_upset(df_all)

# caregiver stress
df_all %>%
  select(id, dep_1:stress_4) %>%
  vis_miss()

df_all %>%
  select(id, dep_1:stress_4) %>%
  gg_miss_upset()
```


## 2.3. Examine Predictors of Missingness 
Based on patterns above, we predict missingness in depression and stress at the 4th measurement (birth), along with EEG measures
```{r}
df_all$miss_dep_4 <- is.na(df_all$dep_4)
df_all$miss_stress_4 <- is.na(df_all$stress_4)

df_all$miss_frontal_sdr <- is.na(df_all$FrontalSync_stdr)
df_all$miss_central_odd <- is.na(df_all$CentralSync_odd)
df_all$miss_par_after <- is.na(df_all$ParietalSync_after)

# setting mom age risk to factor
df_all$mom_age_risk <- as.factor(df_all$mom_age_risk)

# Prediction models
summary(glm(miss_dep_4 ~ dep_1 + dep_2 + dep_3 + stress_1 + stress_2 + stress_3 + mat_med_risk_sum + ibqr_neg + disadv_prenatal + mom_age_risk,
    data = df_all,
    family = binomial))

summary(glm(miss_stress_4 ~ dep_1 + dep_2 + dep_3 + stress_1 + stress_2 + stress_3 + mat_med_risk_sum + ibqr_neg + disadv_prenatal + mom_age_risk,
    data = df_all,
    family = binomial))

summary(glm(miss_frontal_sdr ~ dep_1 + dep_2 + dep_3 + stress_1 + stress_2 + stress_3 + mat_med_risk_sum + ibqr_neg + disadv_prenatal + mom_age_risk,
    data = df_all,
    family = binomial))

summary(glm(miss_central_odd ~ dep_1 + dep_2 + dep_3 + stress_1 + stress_2 + stress_3 + mat_med_risk_sum + ibqr_neg + disadv_prenatal + mom_age_risk,
    data = df_all,
    family = binomial))

summary(glm(miss_par_after ~ dep_1 + dep_2 + dep_3 + stress_1 + stress_2 + stress_3 + mat_med_risk_sum + ibqr_neg + disadv_prenatal + mom_age_risk,
    data = df_all,
    family = binomial))
```

# Imputation overview
We now move to imputation models. These models will be ran in two separate sets. One imputation model will address missingness in repeated measures of depression and stress, to test whether growth models with MI lool similar to the ones estimated with REML.

A second imputation model will address missingness in EEG variables across the trials. Missingness in EEG is relatively large compared to the overall sample, because data collection was only attempted with a subsample (170). So we want to estimate missingness in relation to that, and not the entire enrolled sample (399). These models will be estimated using the exported growth parameters from depression, and average stress, so that we can test the original models.

# 3. Imputation for Depression and Stress Models

## 3.1. Reshape data for running imputation (long format)
Reshaping with usual methods was causing problems with MICE imputation block chains, so I split the dep and sress data then row bind, then merge with everything else.
```{r}
library(tidyr)

t1 <- df_all %>%
  dplyr::select(id, dep_1, stress_1)%>%
  dplyr::mutate(time = 0)%>%
  dplyr::rename(dep=dep_1, stress=stress_1)

t2 <- df_all %>%
  dplyr::select(id, dep_2, stress_2)%>%
  dplyr::mutate(time = 1)%>%
  dplyr::rename(dep=dep_2, stress=stress_2)

t3 <- df_all %>%
  dplyr::select(id, dep_3, stress_3)%>%
  dplyr::mutate(time = 2)%>%
  dplyr::rename(dep=dep_3, stress=stress_3)

t4 <- df_all %>%
  dplyr::select(id, dep_4, stress_4)%>%
  dplyr::mutate(time = 3)%>%
  dplyr::rename(dep=dep_4, stress=stress_4)

long_df_all <- rbind(t1, t2)
long_df_all <- rbind(long_df_all, t3)
long_df_all <- rbind(long_df_all, t4)

# merging with time invariant variables
df_all_tiv <- df_all%>%
  dplyr::select(id:mat_med_risk_sum, disadv_prenatal)
  
long_df_all <- merge(long_df_all, df_all_tiv, by="id", all = T)

long_df_all$time_d <- factor(long_df_all$time,
                          levels = c("0","1","2","3"))

# converting variables to factors
long_df_all$child_sex <- as.factor(long_df_all$child_sex)
long_df_all$mom_age_risk <- as.factor(long_df_all$mom_age_risk)



# making new id column for MICE (the block chains do not work with id as chr structure)
long_df_all <- long_df_all %>%
  dplyr::mutate(id_imp = as.integer(str_extract(id, "\\d+")))

long_df_all<- long_df_all%>%
  dplyr::select(dep:id_imp)%>%
  dplyr::rename(id = id_imp)
```

## 3.2 Descriptives by wave
```{r}
describe.by(long_df_all, group = long_df_all$time_d)

# create centered and quadratic time variables
long_df_all <- long_df_all%>%
  dplyr::mutate(time_c = as.numeric(scale(time, center = T, scale = F)),
                time_c2 = time_c^2)

# removing one participant with only disadvantage variable
long_df_all <- long_df_all%>%
  dplyr::filter(id != 1652)

#inspecting na N
colSums(is.na(long_df_all))
```



## 3.3 Imputation model of both depression and stress as repeated measures

### Combined Predictor Matrix
```{r}
predA <- make.predictorMatrix(long_df_all)

predA["id", ] <- 0       # tells mice id not imputed
predA[,"id"] <- 0  
# Identify cluster variable
predA["dep", "id"] <- -2
predA["stress", "id"] <- -2      # tells mice that id is not a predictor but a grouping variable --> random intercept

```

### Stress Imputation Methods
```{r}
methA <- make.method(long_df_all)

methA["dep"] <- "2l.pmm"  
methA["stress"] <- "2l.pmm"   # continuous repeated measure
methA["child_sex"]      <- "logreg"
methA["mom_age_risk"]      <- "polyreg"
methA["id"]             <- ""
methA["time"]           <- ""

methA
```


### Run combined imputation model
```{r}

impA <- 
  mice:: mice(
  long_df_all,
  m = 40,               # ≥ % missing recommended
  maxit = 50,
  method = methA,
  predictorMatrix = predA,
# blocks = blocks,
  seed = 2025
)

plot(impA)
densityplot(impA)
```


## 4. Depression and Stress growth models

## 4.1. Linear change
```{r}
# depression model
fit_dep <- with(
  impA,
  lmer(
    dep ~ 1 + time +
      (1 | id)
  )
)

pool_dep <- pool(fit_dep)
summary(pool_dep, conf.int = TRUE)


# stress model
fit_st <- with(
  impA,
  lmer(
    stress ~ 1 + time +
      (1 | id)
  )
)

pool_st <- pool(fit_st)
summary(pool_st, conf.int = TRUE)
```
### Results:
Here, neither depression nor stress trajectories sig changed over time

## 4.1. Quadratic change
```{r}
# depression model
fit_depQ <- with(
  impA,
  lmer(
    dep ~ 1 + time_c + time_c2+
      (1  | id)
  )
)

pool_depQ <- pool(fit_depQ)
summary(pool_depQ, conf.int = TRUE)


# stress model
fit_stQ <- with(
  impA,
  lmer(
    stress ~ 1 + time_c + time_c2+
      (1  | id)
  )
)

pool_stQ <- pool(fit_stQ)
summary(pool_stQ, conf.int = TRUE)
```


## 5. Imputation models of delta-beta coupling

## 5.1. Getting the data ready
```{r}
stress_preds <- read.csv(here("data", "stress_preds.csv"), header = T)

all_sync <- merge(all_sync, stress_preds, by="id", all.x = T)
all_sync <- merge(all_sync, ibq, by="id", all.x = T)
all_sync <- merge(all_sync, demo, by="id", all.x = T)
all_sync<- merge(all_sync, ses, by="id", all.x = T)

all_sync <- all_sync%>%
  dplyr::select(-c(n_obs, r.alldata.post.pre.waveleting))

colSums(is.na(all_sync))

# change id column for MICE models
all_sync <- all_sync %>%
  dplyr::mutate(id_imp = as.integer(str_extract(id, "\\d+")))

all_sync<- all_sync%>%
  dplyr::select(trial:id_imp)%>%
  dplyr::rename(id = id_imp)

# separating level-1 and level-2 data
df_lvl2 <- all_sync %>%
  dplyr::select(Dep_int:id)%>%
  dplyr::distinct()


df_lvl1 <- all_sync %>%
  dplyr::select(
    id,
    trial,
    FrontalSync:ParietalSync
  )
#reordering trial for contrast later
df_lvl1$trial <- factor(df_lvl1$trial, ordered = FALSE)
df_lvl1$trial <- relevel(df_lvl1$trial, ref = "odd")
```

### Predictor Matrices
```{r}
pred_lvl2 <- make.predictorMatrix(df_lvl2)
pred_lvl2[, "id"] <- 0
pred_lvl2["id", ] <- 0


pred_lvl1 <- make.predictorMatrix(df_lvl1)

# Declare cluster
pred_lvl1[, "id"] <- -2
pred_lvl1["id", ] <- 0

# Prevent self-prediction
diag(pred_lvl1) <- 0

```

###  Imputation Methods
```{r}
meth_lvl2 <- make.method(df_lvl2)
meth_lvl2["id"] <- ""       # never impute ID
meth_lvl2[] <- "pmm"

meth_lvl1 <- make.method(df_lvl1)
meth_lvl1[c("id", "trial")] <- ""
meth_lvl1["FrontalSync"] <- "2l.pmm"   # continuous repeated measure
meth_lvl1["CentralSync"] <- "2l.pmm" 
meth_lvl1["ParietalSync"] <- "2l.pmm" 


```


### Running the models
```{r}
# level 2 variables (only IBQ being imputed here - everything else has no missingness)
imp_lvl2 <- mice(
  df_lvl2,
  m = 40,
  method = meth_lvl2,
  predictorMatrix = pred_lvl2,
  seed = 123
)
plot(imp_lvl2)
densityplot(imp_lvl2)

# delta-beta coupling variables
imp_lvl1 <- mice(
  df_lvl1,
  m = 40,
  method = meth_lvl1,
  predictorMatrix = pred_lvl1,
  seed = 123
)

plot(imp_lvl1)
densityplot(imp_lvl1)
```

### Merging imputations
```{r}
library(purrr)

lvl1_long <- complete(imp_lvl1, action = "long", include = TRUE)
lvl2_long <- complete(imp_lvl2, action = "long", include = TRUE)

df_imp_long <- lvl1_long %>%
  left_join(
    lvl2_long,
    by = c("id", ".imp")
  )

imp_final <- as.mids(df_imp_long)
```

### Sanity checks
```{r}
# inspecting one dataset
comp1 <- complete(imp_final, 1)

# No missing values left?
colSums(is.na(complete(imp_final, 1)))

```


### Frontal Models
```{r}
FrontalT <- with(
  imp_final,
  lmer(FrontalSync ~ 1 + trial +
              (1 | id) 
            ))
summary(FrontalT)

pool_FrontalT <- pool(FrontalT)
summary(pool_FrontalT, conf.int = TRUE)

# adding covariates
FrontalCOV <-  with(
  imp_final,
  lmer(FrontalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
              (1 | id)
  ))

pool_FrontalCOV <- pool(FrontalCOV)
summary(pool_FrontalCOV, conf.int = TRUE)

# Prenatal Disadvantage
FrontalDis <-  with(
  imp_final,
  lmer(FrontalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         disadv_prenatal*trial +
              (1 | id)
  ))

pool_FrontalDis <- pool(FrontalDis)
summary(pool_FrontalDis, conf.int = TRUE)

# Prenatal Depression
FrontalPren <-  with(
  imp_final,
  lmer(FrontalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
          trial*Dep_int + trial*Dep_QSlope +
              (1 | id)
  ))

pool_FrontalPren <- pool(FrontalPren)
summary(pool_FrontalPren, conf.int = TRUE)

# Prenatal Depression
FrontalStress <-  with(
  imp_final,
  lmer(FrontalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         trial*stress_ave+
         #ibq_neg_m + trial*ibq_neg_m*disadv_prenatal +
              (1 | id)
  ))

pool_FrontalStress <- pool(FrontalStress)
summary(pool_FrontalStress, conf.int = TRUE)

# Prenatal Depression
FrontalTemp <-  with(
  imp_final,
  lmer(FrontalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         trial*ibq_neg_m +
              (1 | id)
  ))

pool_FrontalTemp <- pool(FrontalTemp)
summary(pool_FrontalTemp, conf.int = TRUE)
```

### Central Models
```{r}
CentralT <- with(
  imp_final,
  lmer(CentralSync ~ 1 + trial +
              (1 | id) 
            ))

pool_CentralT<- pool(CentralT)
summary(pool_CentralT, conf.int = TRUE)

# Covariates
CentralCOV <- with(
  imp_final,
  lmer(CentralSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial +
              (1 | id), 
            ))

pool_CentralCOV <- pool(CentralCOV)
summary(pool_CentralCOV, conf.int = TRUE)

# Prenatal Disadvantage
CentralDis <-  with(
  imp_final,
  lmer(CentralSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         disadv_prenatal*trial +
              (1 | id)
  ))

pool_CentralDis <- pool(CentralDis)
summary(pool_CentralDis, conf.int = TRUE)

# Depression
CentralPren <- with(
  imp_final,
  lmer(CentralSync ~ 1 + 
         child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         trial*Dep_int + 
              (1 | id), 
            ))

pool_CentralPren <- pool(CentralPren)
summary(pool_CentralPren, conf.int = TRUE)

# Perceived Stress
CentralStress <- with(
  imp_final,
  lmer(CentralSync ~ 1 + 
         child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
        trial*Dep_int + trial*stress_ave +
              (1 | id), 
            ))

pool_CentralStress <- pool(CentralStress)
summary(pool_CentralStress, conf.int = TRUE)

# IBQ negative affect
CentralTemp <- with(
  imp_final,
  lmer(CentralSync ~ 1 + 
         child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
        trial*Dep_int+ trial*ibq_neg_m +
              (1 | id), 
            ))

pool_CentralTemp <- pool(CentralTemp)
summary(pool_CentralTemp, conf.int = TRUE)
```


### Parietal Models
```{r}
PerT <- with(
  imp_final,
  lmer(ParietalSync ~ 1 + trial +
              (1 | id)
            ))

pool_PerT<- pool(PerT)
summary(pool_PerT, conf.int = TRUE)

PerCOV <- with(
  imp_final,
  lmer(ParietalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial +
              (1 | id), 
            ))

pool_PerCOV <- pool(PerCOV)
summary(pool_PerCOV, conf.int = TRUE)

# Prenatal Disadvantage
ParDis <-  with(
  imp_final,
  lmer(ParietalSync ~ 1 + child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         disadv_prenatal*trial +
              (1 | id)
  ))

pool_ParDis <- pool(ParDis)
summary(pool_ParDis, conf.int = TRUE)

# Depression
ParPren <- with(
  imp_final,
  lmer(ParietalSync ~ 1 + 
         child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         trial*Dep_int + 
              (1 | id), 
            ))

pool_ParPren <- pool(ParPren)
summary(pool_ParPren, conf.int = TRUE)

# Perceived Stress
ParStress <- with(
  imp_final,
  lmer(ParietalSync ~ 1 + 
         child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         trial*stress_ave +
              (1 | id), 
            ))

pool_ParStress <- pool(ParStress)
summary(pool_ParStress, conf.int = TRUE)

# IBQ negative affect
ParTemp <- with(
  imp_final,
  lmer(ParietalSync ~ 1 + 
         child_age_y1_assessment + child_sex + screen_delivery_ga_weeks + trial + 
         trial*ibq_neg_m +
              (1 | id), 
            ))

pool_ParTemp <- pool(ParTemp)
summary(pool_ParTemp, conf.int = TRUE)
```